---
title: "Atlantic Cod Spawning Maps"
output: html_notebook
---


```{r}
source("src/plot_functions.R")
source("src/data.R")

library(data.table)
```

```{r}
#prediction.data = read.csv("data/geo_predictions/cod_preds.csv")
prediction.data = data.table::fread(file="data/geo_predictions/cod_preds.csv")

prediction.data = as.data.frame(prediction.data)
head(prediction.data)
```
```{r}
monthly_environmental_averages = prediction.data %>% 
  dplyr::group_by(year, month) %>% 
  dplyr::summarise(sfc_temp=mean(sfc_temp),
                   sfc_salt=mean(sfc_salt),
                   btm_temp=mean(btm_temp),
                   btm_salt=mean(btm_salt)
                   )

head(monthly_environmental_averages)

```
```{r}
  i=1:dim(monthly_environmental_averages)[[1]]
  monthly_environmental_averages$index = i
  monthly_environmental_averages$date = as.yearmon(
      paste(monthly_environmental_averages$month, monthly_environmental_averages$year, sep="-"), 
      "%m-%Y")
  two.year.rolling=rollmean(monthly_environmental_averages$sfc_temp, 24, na.pad=TRUE)
  monthly_environmental_averages$two.year.rolling = two.year.rolling
  
  options(na.action="na.exclude")
  rolling.lm = lm(two.year.rolling~date, monthly_environmental_averages)
  monthly_environmental_averages$two.year.pred = predict(rolling.lm, data=monthly_environmental_averages)
  summary(rolling.lm)

```
There has been an increase in overall temperatures throughout the NES ecosystem during the handcast timeseries, but the overall increase is small *(+0.017˚C per year)*, though significant *(p<1e-10)*. The linear fit shows that year accounts for only *13%* of the variation seen. _This is expected, as SST is a complex function with many covariates._ 

```{r}
ggplot(monthly_environmental_averages, aes(x=date, y=sfc_temp))+geom_point()+geom_line()+
  geom_line(aes(y=two.year.rolling), color="red", size=1)+
  geom_line(aes( y=two.year.pred), color="blue", size=1)+
  ylim(0, 16)+
  labs(x="Date", y="Surface Temperature (˚C)", title="NES Average Monthly Surface Temperature (1978-2016)")
```
Monthly variation was fairly regular from 1978 through 2010, after which there is a rather distinct rise in SST

```{r}
monthly_position = prediction.data %>% 
  group_by(year, month) %>% 
  summarise(temp=mean(sfc_temp), north=weighted.mean(north, predicted))

monthly_position$date = as.yearmon(
      paste(monthly_position$month, monthly_position$year, sep="-"), 
      "%m-%Y")

monthly.position.model = lm(north~temp, monthly_position)
monthly_position$lm.pred = predict(monthly.position.model, monthly_position)

summary(monthly.position.model)
```
```{r}
ggplot(monthly_position, aes(x=date, y=north))+geom_point()+geom_line()
ggplot(monthly_position)+geom_point(aes(x=temp, y=north, color=month))+geom_line(aes(x=temp, y=lm.pred), color="red")
```
```{r}
annual_position = prediction.data %>% 
  group_by(year) %>% 
  summarise(temp=mean(sfc_temp), north=weighted.mean(north, predicted))

annual.position.model = lm(north~temp, data=annual_position)
annual_position$lm.pred = predict(annual.position.model, annual_position)

summary(annual.position.model)
```
```{r}
ggplot(annual_position)+geom_point(aes(x=temp, y=north, color=year))+geom_line(aes(x=temp, y=lm.pred))+
  labs(x="SST (˚C)", y="Northing", title="Average Habitat Latitude")+xlim(7, 10)

ggplot(annual_position, aes(x=year, y=north))+geom_point()+geom_line()
```


Add a new chunk by clicking the *Insert Chunk* button on the toolbar or by pressing *Cmd+Option+I*.

When you save the notebook, an HTML file containing the code and output will be saved alongside it (click the *Preview* button or press *Cmd+Shift+K* to preview the HTML file). 

The preview shows you a rendered HTML copy of the contents of the editor. Consequently, unlike *Knit*, *Preview* does not run any R code chunks. Instead, the output of the chunk when it was last run in the editor is displayed.

